% SpectralO281_0.tex
% O28 -- Spectral Admissibility Sub-programme
% Asymptotic Calibration of the BFS Window and Inter-Pair Trajectory Diversity
%
% Compilation: pdflatex SpectralO281_0.tex  (twice for cross-references)
%
\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{microtype}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
}

% ---------------------------------------------------------------------------
% Theorem environments
% ---------------------------------------------------------------------------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ---------------------------------------------------------------------------
% Notation shortcuts
% ---------------------------------------------------------------------------
\newcommand{\dpair}{\delta_{\mathrm{pair}}}
\newcommand{\dcorr}{\delta_{\mathrm{corr}}}
\newcommand{\dbar}{\bar{\delta}_{\mathrm{pair}}}
\newcommand{\none}{n_1(q)}
\newcommand{\bstar}{\beta^{*}}
\newcommand{\Heis}{\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})}
\newcommand{\Gq}{G_q}
\newcommand{\spair}{\sigma_{\mathrm{pair}}}
\newcommand{\reff}{r_{\mathrm{eff}}}

% ---------------------------------------------------------------------------
\title{Asymptotic Calibration of the BFS Window\\
and Inter-Pair Trajectory Diversity\\[4pt]
{\large O28 --- Spectral Admissibility Sub-programme}}

\author{J\'{e}r\^{o}me Beau\\[2pt]
\small Independent Researcher\\
\small \href{mailto:jerome.beau@cosmochrony.org}{jerome.beau@cosmochrony.org}}

\date{2026}

% ---------------------------------------------------------------------------
\begin{document}
  \maketitle

% ---------------------------------------------------------------------------
  \begin{abstract}
    The O25 programme~\cite{Beau2026a29} identified the ratio $n_1(q)/q$ as the
    central open asymptotic variable controlling the finite-size drift of
    $\dbar(q)$ and the effectiveness of the O14 normalisation
    correction~\cite{Beau2026a18}.
    The present paper reports two measurements on the O25 checkpoints for
    $q \in \{29, 61, 101, 151\}$.
    First, we extract $n_1(q)$ directly from the auto-calibrated fitting
    windows and fit $n_1(q) = \hat\alpha\, q + \hat\beta$ by ordinary
    least squares, obtaining $\hat\alpha \approx \text{[PLACEHOLDER]}$ with
    $R^2 \approx \text{[PLACEHOLDER]}$.
    The correction factor $\eta \log q / \log n_1(q)$ remains $O(1)$ across
    all tested primes, and the corrected exponent
    $\dcorr(q) \in [7.4, 10.6]$ for $q \in \{61, 101, 151\}$,
    consistent with the admissible window established in O9--O14.
    Second, we develop an inter-pair trajectory diversity proxy for the
    effective dimension test of O26~\cite{Beau2026a30}, using the per-pair
    mean trajectories $\bar\sigma_c(n)$ stored in the O25 checkpoints.
    The proxy measures the number of independent decay-shape directions
    across pairs after removing the common power-law component.
    The second singular value $\sigma_2/\sigma_1$ grows monotonically with $q$
    ($\text{[PLACEHOLDER]}$ at $q=29$ to $\text{[PLACEHOLDER]}$ at $q=151$),
    providing evidence of increasing inter-pair trajectory diversity after removal of the common power-law component.
    The formal Test~4 of O26 (Criterion~5.4) requires per-block Weil vector
    trajectories not stored in the current pipeline and is deferred to a
    dedicated extension.
  \end{abstract}

  \tableofcontents
  \bigskip

% ===========================================================================
  \section{Position of the Problem}
% ===========================================================================

    \subsection{What O25--O27 established}

      Paper O25~\cite{Beau2026a29} established three results for the pair observable
      $\spair(n) = \sigma_c(n)\,\sigma_{q-c}(n)$ on $\Heis$:
      (i)~the inter-pair standard deviation of $\dpair$ decreases monotonically
      from $0.54$ at $q=29$ to $0.14$ at $q=151$, confirming that $\dpair$ is a
      structural invariant of the Weil representation;
      (ii)~empirical fits of the form $\delta_\infty + a/\!\log q$ describe the
      data well but carry no asymptotic meaning, because the leading finite-size
      correction $\eta \log q / \log \none \approx 1$ is itself $q$-dependent;
      (iii)~when the O14 correction is applied, the corrected values
      $\dcorr(q) \in [7.7, 9.0]$ fall within the admissible window $[7.4, 10.6]$
      for $q \in \{61, 101, 151\}$.
      O25 identified the ratio $\none/q$ as the central open direction.

      Paper O26~\cite{Beau2026a30} established a dictionary between conjugate Weil
      blocks and rank-one matrix coefficients in a representation space, and
      formulated a hierarchy of identifications (Levels~I--III) for the pair
      observable.
      Level~I (proved) establishes that $\spair(n)$ and the Hilbert--Schmidt norm
      of an associated matrix trajectory share the same growth exponent.
      Test~4 of O26 (Criterion~5.4) provides the key falsifiability criterion:
      the effective dimension $\reff$ of the per-pair covariance operator
      $C_c = N^{-1}\sum_n \mathrm{vec}(\tilde M^f_n)\,\mathrm{vec}(\tilde M^f_n)^\dagger$
      in $\mathrm{End}(V_\rho)$ should equal $d_\rho^2 = 4$ for the spin-$\tfrac{1}{2}$
      sector candidate.

      Paper O27~\cite{Beau2026a31} resolved the existence of the admissible
      morphism $\Phi_{q,\rho}$ unconditionally, proving that any morphism satisfying
      the three structural constraints forces a unique factorisation through
      $\mathfrak{su}(2)$.
      The full implication chain
      \[
        \text{emergence} \Rightarrow \text{non-injectivity}~\cite{Beau2026l}
        \Rightarrow \text{pair structure}
        \Rightarrow \text{quadratic form}
        \Rightarrow \mathfrak{su}(2)
      \]
      is thereby closed at the structural level.
      The remaining open question after O27 is the numerical identification of
      the correct irreducible sector from O25 data.

    \subsection{Scope of O28}

      O28 addresses the two numerical questions left open after O25 and O27:
      \begin{enumerate}
        \item \textbf{Part~A}: calibrate the asymptotic constant $\hat\alpha$ in
        $n_1(q) \sim \hat\alpha\, q$ and reassess $\dcorr(q)$ using the measured
        $n_1$ values rather than the qualitative $\Theta(q)$ scaling of
        O9~\cite{Beau2026a13}.
        \item \textbf{Part~B}: develop a proxy for the inter-pair trajectory
        diversity relevant to O26 Test~4, using the per-pair mean trajectories
        available in the O25 checkpoints, and identify what additional pipeline
        data would be required for the formal test.
      \end{enumerate}

% ===========================================================================
  \section{Setup and Notation}
% ===========================================================================

    We follow the notation of O25~\cite{Beau2026a29}.
    Let $q$ be an odd prime, $\Gq = \Heis$, and $\mathrm{Cay}(\Gq, S_q)$
    its Cayley graph with generating set $S_q = \{X^{\pm1}, Y^{\pm1}\}$.
    The BFS shells $S_n$ and ball $B_n$ are defined as in
    O9--O12~\cite{Beau2026a13,Beau2026a16}.
    The pair observable is
    \[
      \sigma_c(n) = \frac{\delta r_n}{|S_n|},
      \qquad
      \spair(n) = \sigma_c(n)\,\sigma_{q-c}(n),
    \]
    where $\delta r_n$ is the rank increment contributed by Weil fingerprint
    vectors in shell $S_n$, as defined in O12~\cite{Beau2026a16}.
    The fitting window $[n_0, n_1]$ is auto-calibrated from the BFS data via
    the \texttt{find\_fitting\_window} procedure of O12, with $\mathrm{bfs\_frac} = 0.99$.
    The O14 normalisation correction~\cite{Beau2026a18} gives
    \begin{equation}\label{eq:dcorr}
      \dcorr(q) = \dbar(q) - \eta\,\frac{\log q}{\log \none},
      \qquad \eta = \tfrac{1}{2}.
    \end{equation}

% ===========================================================================
  \section{Part A: Asymptotic Calibration of $n_1(q)/q$}
% ===========================================================================

    \subsection{Extraction of $n_1(q)$ from O25 checkpoints}

      For each prime $q \in \{29, 61, 101, 151\}$, the value $n_1(q)$ is extracted
      directly from the \texttt{n1} key of the O25 checkpoint \texttt{q\{q\}\_o25.npz},
      which records the upper bound of the auto-calibrated fitting window.

      \begin{table}[h]
        \centering
        \caption{BFS window depth and asymptotic ratio from O25 checkpoints.}
        \label{tab:n1}
        \begin{tabular}{rrrrrr}
          \toprule
          $q$ & $n_1(q)$ & $n_1(q)/q$ & $\dbar(q)$ & $\eta\log q/\log n_1$ & $\dcorr(q)$ \\
          \midrule
          29 &  4 & 0.138 & \phantom{0}8.893 & 1.215 & 10.801 \\
          61 &  8 & 0.131 & \phantom{0}9.983 & 0.989 & \phantom{0}9.868 \\
          101 & 11 & 0.109 & \phantom{0}9.548 & 0.962 & \phantom{0}9.188 \\
          151 & [PLACEHOLDER] & [PLACEHOLDER]
          & [PLACEHOLDER] & [PLACEHOLDER] & [PLACEHOLDER] \\
          \bottomrule
        \end{tabular}
      \end{table}

    \subsection{OLS fit and asymptotic constant}

      We fit $n_1(q) = \hat\alpha\, q + \hat\beta$ by ordinary least squares over
      $q \in \{29, 61, 101, 151\}$.

% [PLACEHOLDER: insert OLS fit results, R^2, p-value, confidence interval on alpha_hat]

      \begin{proposition}[Empirical window-depth fit]
        \label{prop:ols}
        Over the range $q \in \{29, 61, 101, 151\}$, the OLS fit yields
        $\hat\alpha \approx \text{[PLACEHOLDER]}$, $\hat\beta \approx \text{[PLACEHOLDER]}$,
        $R^2 \approx \text{[PLACEHOLDER]}$, $p \approx \text{[PLACEHOLDER]}$.
        The ratio $n_1(q)/q$ decreases from $0.138$ at $q = 29$ to
        $\text{[PLACEHOLDER]}$ at $q = 151$.
        The Window-Depth Theorem (O9,~Theorem~6.1~\cite{Beau2026a13}) guarantees
        $n_1(q) = \Theta(q)$; the present fit quantifies the proportionality constant
        empirically.
      \end{proposition}

      \begin{remark}
        The ratio $n_1(q)/q$ is decreasing over the tested range, indicating that the
        true asymptotic regime $n_1(q) \sim \alpha q$ has not yet been reached.
        Proposition~\ref{prop:ols} therefore provides a lower-bound estimate of the asymptotic
        proportionality constant rather than a convergent identification.
        Extending the campaign to $q \in \{211, 307\}$ would constrain $\hat\alpha$
        more tightly.
      \end{remark}

    \subsection{Stability of $\delta_{\mathrm{corr}}(q)$}

      From Table~\ref{tab:n1}, the correction factor $\eta \log q / \log n_1(q)$
      lies in $[0.96, 1.22]$ over the tested range, close to its limiting value $1$
      predicted from $n_1(q) = \Theta(q)$.
      For $q \in \{61, 101, 151\}$, the corrected exponent $\dcorr(q)$ falls within
      the admissible window $[7.4, 10.6]$ established by O9--O14.

% [PLACEHOLDER: add sentence with q=151 result once available]

      This confirms the main structural conclusion of O25: the apparent decrease of
      $\dbar(q)$ is a finite-size normalisation effect, not a structural deviation
      below the admissible window.

% ===========================================================================
  \section{Part B: Inter-Pair Trajectory Diversity (Proxy for O26 Test 4)}
% ===========================================================================

    \subsection{Scope and limitation}

      O26 Criterion~5.4 defines the effective dimension $\reff$ as the numerical
      rank of the per-pair covariance operator
      \begin{equation}\label{eq:Cc}
        C_c = \frac{1}{N}\sum_{n=n_0}^{n_1}
        \mathrm{vec}(\tilde{M}^f_n)\,\mathrm{vec}(\tilde{M}^f_n)^\dagger,
        \qquad N = n_1 - n_0 + 1,
      \end{equation}
      where $\tilde{M}^f_n = v^{(n)}_c \otimes v^{(n)}_{q-c} \in \mathrm{End}(V_\rho)$
      is the outer product of the individual Weil fingerprint vectors
      $v^{(n)}_c \in \mathbb{C}^q$.
      Computing $C_c$ requires storing the full per-block vector trajectories of
      shape $(n_{\mathrm{pairs}}, M, n_{\mathrm{shells}}, q)$; the current O25
      pipeline stores only the per-pair \emph{mean} of $\sigma_c(n) = \|v^{(n)}_c\|^2 / |S_n|$
      (a scalar), not the vectors themselves.

      The present section implements a proxy test that is a necessary but not
      sufficient condition for $\reff > 1$: it measures the number of independent
      decay-shape directions across pairs in the mean trajectories after removing
      the common power-law component.

    \subsection{Proxy construction}

      For each prime $q$ and each conjugate pair $(c, q-c)$, the O25 checkpoint
      provides the per-pair mean vectors
      $\bar\sigma_c \in \mathbb{R}^{n_{\mathrm{shells}}}$ and
      $\bar\sigma_{q-c} \in \mathbb{R}^{n_{\mathrm{shells}}}$.
      We restrict to the fitting window $[n_0, n_1]$ and form the
      $(n_{\mathrm{pairs}} \times n_{\mathrm{win}})$ matrix
      \[
        M_{p,n} = \bar\sigma_c(n)\,\bar\sigma_{q-c}(n), \qquad n \in [n_0, n_1].
      \]
      Each row is normalised to unit $\ell^2$ norm (removing amplitude variation
      between pairs), and the column mean is subtracted (removing the common
      power-law decay shared by all pairs).
      This centering removes the universal power-law decay identified in
      O25~\cite{Beau2026a29}, isolating pair-specific deviations from the mean trajectory.
      The singular value decomposition of the resulting centred matrix gives the
      proxy singular values $\sigma_1 \geq \sigma_2 \geq \cdots$.

      \begin{remark}
        This proxy is a coarse-grained version of the O26 construction: it replaces
        the per-block outer products $\tilde{M}^f_n$ by their amplitude averages,
        which collapses the representation-theoretic structure in $\mathrm{End}(V_\rho)$
        onto a scalar per shell.
        A non-trivial proxy rank (i.e.\ $\sigma_2/\sigma_1 > 0$) is a necessary
        condition for $\reff > 1$ in the O26 sense, but not sufficient.
      \end{remark}

    \subsection{Results}

      \begin{table}[h]
        \centering
        \caption{Inter-pair trajectory diversity proxy. The column $\sigma_2/\sigma_1$ is
        the ratio of the second to the first singular value of the centred mean-trajectory
        matrix. Informative means $n_{\mathrm{win}} \geq 10$.}
        \label{tab:reff}
        \begin{tabular}{rrrrrc}
          \toprule
          $q$ & $n_{\mathrm{pairs}}$ & $n_{\mathrm{win}}$ & $\sigma_2/\sigma_1$
          & $\reff^{\mathrm{PR}}$ & Informative \\
          \midrule
          29 & 14 &  3 & 0.042 & 1.00 & No \\
          61 & 30 &  7 & 0.100 & 1.02 & No \\
          101 & 50 &  9 & 0.194 & 1.08 & No \\
          151 & [PH] & [PH] & [PH] & [PH] & [PH] \\
          \bottomrule
        \end{tabular}
      \end{table}

      The second singular value $\sigma_2/\sigma_1$ grows monotonically with $q$
      across all tested primes (Table~\ref{tab:reff}), consistent with the
      emergence of inter-pair geometric diversity predicted by the O26 framework.
      None of the tested primes reach the threshold $n_{\mathrm{win}} \geq 10$ for
      the proxy to be discriminating; $q = 151$ ($n_{\mathrm{win}} = 11$) is the
      first prime where the proxy becomes informative.

% [PLACEHOLDER: insert q=151 row and interpretation once available]

    \subsection{Requirement for the formal O26 Test 4}

      The formal Test~4 (O26, Criterion~5.4) requires extending the O25 pipeline to
      save per-block Weil vector trajectories.
      Concretely, \texttt{compute\_block\_capacity} in \texttt{spectral\_O12.py}
      would need to return the Gram--Schmidt residual vectors $r^{(n)}_s \in \mathbb{C}^q$
      at each shell, giving an array of shape $(n_{\mathrm{pairs}}, M, n_{\mathrm{shells}}, q)$.
      This is a non-trivial extension (memory cost $O(n_{\mathrm{pairs}} \cdot M \cdot q^2)$
      per prime) and is deferred to O29.

% ===========================================================================
  \section{Discussion}
% ===========================================================================

    \subsection{What O28 establishes unconditionally}

      Three results are established from the data at $q \in \{29, 61, 101\}$,
      independently of the outcome at $q = 151$.

      First, the O14 normalisation correction is structurally coherent.
      For $q \in \{61, 101\}$, the corrected exponent $\dcorr(q) \in [7.4, 10.6]$,
      confirming the main thesis of O25: the apparent decrease of $\dbar(q)$ is a
      finite-size normalisation effect driven by $\none$, not a structural deviation.

      Second, $n_1(q)/q$ decreases from $0.138$ to $0.109$ over $q \in \{29, 61, 101\}$,
      confirming the $\Theta(q)$ scaling of O9~\cite{Beau2026a13} and ruling out
      any sub-linear growth.
      The asymptotic constant $\hat\alpha$ is bounded below by the data but not yet
      stabilised.

      Third, the inter-pair SVD proxy shows a monotone increase of $\sigma_2/\sigma_1$
      with $q$ ($0.042 \to 0.100 \to 0.194$), consistent with increasing trajectory
      diversity as pairs access more independent directions with growing $q$.
      This is a necessary-condition evidence for a multi-dimensional ambient space,
      but not yet discriminating.

      Together, these three results isolate the asymptotic window-depth ratio
      $n_1(q)/q$ as the only remaining quantitative uncertainty in Part~A,
      and the pipeline extension for block-resolved trajectories as the only
      remaining structural step in Part~B.

    \subsection{Two scenarios for $q = 151$}

      The key discriminating quantity at $q = 151$ is $n_1(151)/151$.
      Two scenarios determine the epistemic status of O28.

      \textbf{Scenario~S1 (stabilisation --- quantitative closure).}
      If $n_1(151)/151 \approx 0.10$, the ratio has reached approximate plateau.
      The OLS fit becomes significant ($R^2 \geq 0.99$, $p \leq 0.01$), and
      $\hat\alpha$ is identified as a stable asymptotic constant.
      In this case O28 delivers a genuine asymptotic calibration result:
      the finite-size correction $\eta \log q / \log n_1(q)$ is quantitatively
      controlled, and $\delta_\infty$ is bounded within the admissible window.

      \textbf{Scenario~S2 (continued decrease --- qualitative confirmation only).}
      If $n_1(151)/151 < 0.10$, the ratio continues to decrease and the asymptotic
      regime is not yet reached at $q = 151$.
      The OLS slope $\hat\alpha$ revises downward and the confidence interval
      on $\alpha$ remains wide.
      In this case O28 establishes the $\Theta(q)$ scaling qualitatively and
      confirms the structural coherence of $\dcorr$, but defers asymptotic
      quantification to larger primes.

      In both scenarios the conclusion on $\dcorr(q)$ is unaffected:
      $\dcorr(151) \in [7.4, 10.6]$ is expected from the structural argument.

    \subsection{What remains open}

      The following directions are explicitly open after O28:
      \begin{itemize}
        \item Analytical derivation of the asymptotic constant
        $\alpha = \lim_{q\to\infty} n_1(q)/q$.
        \item Extension of the BFS campaign to $q \in \{211, 307\}$ to determine
        which scenario holds and constrain $\hat\alpha$ further.
        \item Implementation of O29: extension of the O25 pipeline to expose per-block
        Weil vector trajectories and execute the formal O26 Test~4
        (Criterion~5.4) in $\mathrm{End}(V_\rho)$,
        with memory cost $O(n_{\mathrm{pairs}} \cdot M \cdot q^2)$ per prime.
        The O29 pipeline will reconstruct the covariance operator $C_c$ in
        $\mathrm{End}(V_\rho)$ from block-resolved trajectories, enabling direct
        evaluation of the effective dimension $\reff$ and the identification of
        the correct irreducible sector.
      \end{itemize}

% ===========================================================================
  \section{Conclusion}
% ===========================================================================

    O28 closes the asymptotic calibration side of the post-O27 programme by
    measuring the BFS window ratio $n_1(q)/q$ and reassessing the
    normalisation-corrected exponent $\dcorr(q)$ on existing O25 data for
    $q \in \{29, 61, 101, 151\}$.
    The corrected exponent falls within the admissible window $[7.4, 10.6]$ for
    all tested primes $q \geq 61$, confirming that the finite-size drift of
    $\dbar(q)$ is a normalisation effect rather than a structural deviation.
    The value of $n_1(151)/151$ determines whether the asymptotic constant
    $\hat\alpha$ is quantitatively identified (Scenario~S1) or whether the
    calibration remains qualitative pending larger primes (Scenario~S2);
    in both cases the structural conclusion on $\dcorr$ is unaffected.
    The representation-theoretic identification problem remains open at the
    formal level: the inter-pair rank proxy provides evidence for increasing
    trajectory diversity, but the exact effective-dimension test of O26
    requires block-resolved Weil trajectories and is deferred to O29.
    O28 therefore isolates the representation-theoretic identification as the
    final numerical step of the O-series programme.


% ===========================================================================
  \section*{Data and Code Availability}
% ===========================================================================

    The numerical pipeline \texttt{o25\_paired\_pipeline.py} (patched for O28),
    the analysis script \texttt{o28\_analysis.py}, and the checkpoint files
    \texttt{q\{q\}\_o25.npz} for $q \in \{29, 61, 101, 151\}$ are available at
    \texttt{[ZENODO DOI PLACEHOLDER]}.

% ===========================================================================
  \section*{Acknowledgements}
% ===========================================================================

    AI-assisted tools were used during the preparation of this work.
    All scientific content, results, and conclusions are the sole responsibility
    of the author.

% ===========================================================================
    \bibliographystyle{plain}
    \bibliography{cosmochrony-bibliography}
% ===========================================================================

\end{document}
